---
title: "Assignment: Bara et al. 2007"
author: "Text as Data"
date: "9/24/2020"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "")
```

## Bara dictionary

Read in the corpus of texts from the Bara et al. paper (hint: you can find it in the `data` folder of `/cloud/project`). For comparison to the paper, you'll want `corpus_bara_speaker`.

Now read in our translation of their dictionary `bara-et-al.ykd` (originally from different software).

```{r}

```

## Replication Time!

### Table 1

Can you replicate Table 1? How close can you get? How do your answers differ, (if they do)?

```{r}

```


### Table 2

Can you replicate Table 2? What is the largest divergence? Qualitatively speaking does your analysis agree?

```{r}

```

This is exactly the kind of summary construction and analysis we need to be comfortable doing. It should take you a few lines of R, so if you are writing reams of code, maybe seek advice on you R! (I'm happy to help if you're stuck)

### `Keyness'

Following the paper, we can use the `vote` document variable to distinguish groups of speakers who subsequently voted for and those who voted against the legislation.

Create a `dfm` for the corpus, then restrict it to a subset of vocabulary of substantive interest, e.g. you might choose explicitly gendered language, e.g. he, him, she, her, or maybe occupational language. This is your choice - be as wide or as narrow as you find interesting. 

Now use `textstat_keyness` to compare vocabulary usage between these two groups. (This function is a variation on the log ratio scores we have been seeing in lecture.) Very briefly, describe your results. 

Note that you will want to assign `groups` to the `dfm` when measuring keyness. (The help pages will help here!)

```{r}

```

## Handing it all in

This time around, just compile this document with code and answers and upload it to moodle. Specifically, I'll be looking for a single knitted pdf, Word, or html file. I think that's probably fancy enough, while we get used to the cloud business. 
